Figure two

### 2025-12-17 dataPrep.Rmd
updated code and figure but need to review fig2 and update exposure fig2 heatmap

@@

@@

```{r count eurobarometer}

library(tidyverse)
surveys <- read.csv("output/surveyClassifYear.csv")

surveys <- surveys %>% 
  filter(Inclusion == 1)

surveys <- surveys %>% 
  filter(!is.na(iso3))

surveys <- surveys %>% 
  filter(!(iso3 == "FRA" & continent == "Seven seas (open ocean)")) %>% 
  mutate(item_id = paste0(iso3, id, year)) %>%                      # unique survey item ID
  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category

eurobarometer_surveys <- surveyClassification %>%
  filter(grepl("Eurobarometer", SurveyName, ignore.case = TRUE))

nrow(eurobarometer_surveys)

>>>>

# count separately for eurobarometer and non-eurobarometer
surveyCountries_long_clean_t_euro <- surveyCountries_long_clean_t %>% 
  mutate(eurobarometerSource = ifelse(grepl("Eurobarometer", source, ignore.case = TRUE), 1, 0))

@@

# create a barometer-split dataset

surveyCountries_long_clean_t_euro <- surveyCountries_long_clean_t_euro

surveys_by_country_euro <- surveyCountries_long_clean_t_euro %>%
  filter(Inclusion == 1) %>% 
  group_by(iso3, eurobarometerSource, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

summaryDataset_euro <- merge(summaryDataset2 %>% 
                               select(-n_surveys) %>% 
                               crossing(eurobarometerSource = c(0, 1)),
                             surveys_by_country_euro,
                             by = c("iso3", "eurobarometerSource"), all.x = T, all.y = T)

summaryDataset_euro$n_surveys <- ifelse(is.na(summaryDataset_euro$n_surveys), 0, summaryDataset_euro$n_surveys)

write.csv(summaryDataset_euro, "output/datasetPlots_eurobarometer.csv", row.names = F)

@@

### 2025-12-17 draftFigure1.Rmd
updated code and figure but need to review fig2 and update exposure fig2 heatmap

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "income Group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

@@

```{r panels with eurobarometer highlight}
library(tidyverse)
dataset <- read.csv("output/datasetPlots.csv")

>>>>

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

@@

```{r panels with eurobarometer highlight}
# globally there are 2390 survey items from eurobarometer out of 3140 total

library(tidyverse)
dataset <- read.csv("output/datasetPlots_eurobarometer.csv")

dataset <- dataset %>% 
  mutate(isDiff = ifelse(eurobarometerSource == 1, "Eurobarometer", "Rest"))

summarySurveysInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(Income.group))

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

pdf("plots/Figure1_surveys_by_income_eurob.pdf", width = 10)
p2
dev.off()

p4 <- summarySurveysExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

pdf("plots/Figure1_surveys_by_exposure_eurob.pdf", width = 10)
p4
dev.off()

### 2025-12-17 draftFigure2.Rmd
edit survey classif with more items
more surveys need to add whether barometer
updated code and figure but need to review fig2 and update exposure fig2 heatmap

## subset survey without Flashbarometer
surveys_sub <- surveys[which(surveys$source != "Flash Eurobarometer 360: Attitudes of Europeans towards air quality"),]

@@

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_survey_barplot_share.pdf", width = 12, height = 10)
ggplot(surveys_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            hjust = -0.3, size = 9) +
  labs(y = NULL, x = "Number of survey items covering each category") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  xlim(0, 1850) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 22),
    axis.text.x = element_text(size = 22),
    axis.title = element_text(size = 23)
    )
dev.off()

@@

# For each paper, how many distinct countries are there in total,
# and how many of those are in each income group?
paper_country_weights <- papers %>%
  distinct(ID, country, Income.group) %>%  # ignore category here
  group_by(ID) %>%
  mutate(n_countries_total = n_distinct(country)) %>%
  group_by(ID, Income.group, n_countries_total) %>%
  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
  mutate(country_weight = n_countries_group / n_countries_total)
# Example: if a paper has 2 countries (1 low, 1 high) then
# country_weight = 0.5 for low income and 0.5 for high income.

## 2. Attach weights back to the (paper x income group x category) rows ------

papers_weighted <- papers %>%
  left_join(
    paper_country_weights %>%
      dplyr::select(ID, Income.group, country_weight),
    by = c("ID", "Income.group")
  )

@@

topics_data <- papers_weighted %>%
  filter(!is.na(Income.group), !is.na(category)) %>%
  group_by(Income.group, category) %>%
  summarise(
    weighted_n = sum(country_weight, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(Topic = category) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  ) %>%
  group_by(Income.group) %>%
  mutate(
    share = weighted_n / sum(weighted_n),
    total = sum(weighted_n)        # weighted total "papers" per income group
  ) %>%
  ungroup()

>>>>

## subset survey without Flashbarometer
# surveys_sub <- surveys[which(surveys$source != "Flash Eurobarometer 360: Attitudes of Europeans towards air quality"),]

@@

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_survey_barplot_share.pdf", width = 12, height = 10)
ggplot(surveys_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            hjust = -0.3, size = 9) +
  labs(y = NULL, x = "Number of survey items covering each category") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  xlim(0, 2100) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 22),
    axis.text.x = element_text(size = 22),
    axis.title = element_text(size = 23)
    )
dev.off()

@@

## Alternative approach: also account for multiple categories per paper -----
paper_country_weights <- papers %>%
  distinct(ID, country, Income.group) %>%
  group_by(ID) %>%
  mutate(n_countries_total = n_distinct(country)) %>%
  group_by(ID, Income.group, n_countries_total) %>%
  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
  mutate(country_weight = n_countries_group / n_countries_total) %>%
  dplyr::select(ID, Income.group, country_weight)

papers_pic <- papers %>%
  filter(!is.na(Income.group), !is.na(category)) %>%
  distinct(ID, Income.group, category) %>%
  group_by(ID) %>%
  mutate(n_cat = n_distinct(category)) %>%
  ungroup() %>%
  left_join(paper_country_weights, by = c("ID", "Income.group")) %>%
  mutate(weight = country_weight / n_cat)

topics_data <- papers_pic %>%
  group_by(Income.group, category) %>%
  summarise(weighted_n = sum(weight, na.rm = TRUE), .groups = "drop") %>%
  rename(Topic = category) %>%
  group_by(Income.group) %>%
  mutate(
    share = weighted_n / sum(weighted_n),
    total = sum(weighted_n)
  ) %>%
  ungroup() %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  )

@@

# topics_data <- papers_weighted %>%
#   filter(!is.na(Income.group), !is.na(category)) %>%
#   group_by(Income.group, category) %>%
#   summarise(
#     weighted_n = sum(country_weight, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   rename(Topic = category) %>%
#   mutate(
#     Income.group = factor(
#       Income.group,
#       levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
#     ),
#     Topic = factor(
#       Topic,
#       levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
#     )
#   ) %>%
#   group_by(Income.group) %>%
#   mutate(
#     share = weighted_n / sum(weighted_n),
#     total = sum(weighted_n)        # weighted total "papers" per income group
#   ) %>%
#   ungroup()
